---
title: "STAT 7365 
HW_01 (Total Points: 50)"
output: html_notebook
---

The first step in this project is to install base R version 4.3.2

The second step is to download and install RStudio-2023.12.1-402

load the required library (Packages) eg. ggplot2, methods, stats, utils,

To see a list of available datasets in R

```{r}
library(datasets)
```

To display the datasets present in r

```{r}
data()
```

```{r}
data(iris)
```

### Description

The iris data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica. iris is a data frame with 150 cases (rows) and 5 variables (columns) named Sepal.Length, Sepal.Width, Petal.Length, Petal.Width, and Species.

1.  Read the iris data_set from R package data_sets and produce the following graphs.

To read the iris dataset

```{r}
data(iris)
```

To view the structure of the dataset

```{r}
str(iris)
```

Display the first few rows of the dataset

```{r}
head(iris)
```

To view the entire dataset

```{r}
iris
```

A)  Create a scatter plot of Sepal Width and Sepal Length.

```{r}
plot(iris$Sepal.Length, iris$Sepal.Width, main = "Scatter Plot", xlab = "Sepal Length", ylab = "Sepal Width", col = iris$Species)
legend("topright", legend = levels(iris$Species), col = 1:3, pch = 1)
```

the first line of code creates a scatter plot of Sepal.Length against Sepal.Width #with different colors for each species. #The second line of code adds a legend to the plot.

B)  Create a bar plot of Species.

Count of all occurrences of each species

```{r}
species_counts <- table(iris$Species)
```

To create a bar plot of the species

```{r}
barplot(species_counts, main = "Barplot of Species Counts in Iris Dataset", xlab = "Species", ylab = "Count", col = c("red", "green", "blue"), legend.text = TRUE)
```

C)  Create a box plot of Species by Sepal Width (X-axis - Species, Y-axis sepal width).

```{r}
boxplot(Sepal.Width ~ Species, data = iris, main = "Boxplot of Sepal Width by Species", xlab = "Species", ylab = "Sepal Width", col = c("red", "green", "blue"))
```

D)  Create a histogram of Sepal Length. Try using fill= Species statement.

```{r}
hist(iris$Sepal.Length, col = c("red", "green", "blue"),
     main = "Histogram of Sepal Length by Species",
     xlab = "Sepal Length", ylab = "Frequency",
     border = "black",
     legend = TRUE,
     labels = levels(iris$Species), density = 20)
```

E)  Create a histogram of Sepal Length by setting the number of bins=20.

```{r}
hist(iris$Sepal.Length, breaks = 20, main = "Histogram of Sepal Length (20 Bins)", xlab = "Sepal Length", ylab = "Frequency", border = "black")
```

F)  Create a KDE plot of Sepal Length. Use color= Species.

To create a KDE (Kernel Density Estimation) plot of Sepal Length in R with

colors specified by Species,

I will use the geom_density() function from the ggplot2 package.

```{r}
library(ggplot2)

# Create the KDE plot
ggplot(iris, aes(x = Sepal.Length, fill = Species)) +
  geom_density(alpha = 0.5) +
  labs(title = "Stacked KDE Plot of Sepal Length by Species",
       x = "Sepal Length (cm)",
       y = "Density") +
  theme_minimal()
```

G)  Create s stacked KDE plot of Sepal Length.

To create a stacked Kernel Density Estimation (KDE) plot of Sepal Length in R,

I will use the geom_density() function from the ggplot2 package. To make it stacked,

I will use the fill aesthetic to differentiate the distributions.

```{r}
ggplot(iris, aes(x = Sepal.Length, fill = Species)) + geom_density(alpha = 0.5) + labs(title = "Stacked KDE Plot of Sepal Length", x = "Sepal Length", y = "Density")
```

2.  Read the Chick Weight data from R package datasets and produce the following graphs.

library(datasets) data() data(ChickWeight) ChickWeight

Description The ChickWeight data frame has 578 rows and 4 columns from an experiment on \# the effect of diet on early growth of chicks. \# An object of class c("nfnGroupedData", "nfGroupedData", "groupedData", "data.frame") containing the following columns:

weight: a numeric vector giving the body weight of the chick (gm).

Time: a numeric vector giving the number of days since birth when the measurement was made.

Chick: an ordered factor with levels 18 \< ... \< 48 giving a unique identifier for the chick. The ordering of the levels groups chicks on the same diet together and orders them according to their final weight (lightest to heaviest) within diet.

Diet: a factor with levels 1, ..., 4 indicating which experimental diet the chick received.

Details

The body weights of the chicks were measured at birth and every second day thereafter until day 20. They were also measured on day 21. There were four groups on chicks on different protein diets.

This dataset was originally part of package nlme, and that has methods (including for [, as.data.frame, plot and print) for its grouped-data classes.

A)  Create a line plot for weight of all chicks (Hint: x= Time, y= weight, color=chicks).

To create a line plot for the weight of all chicks over time in R,

you can use the ggplot2 package. I'll provide an example assuming you have a

dataset with columns named Time, Weight, and Chicks. Make sure to replace these

with the actual column names in your dataset.

```{r}
ggplot(ChickWeight, aes(x = Time, y = weight, color = Chick)) + geom_line() + labs(title = "Line Plot of Weight over Time for Chicks", x = "Time", y = "Weight", color = "Chick")
```

B)  Create a grid of line plots of weight of chicks by diets.

(Hint: create 4 plots in a single grid with x= Time, Y= Weight)

```{r}
ggplot(ChickWeight, aes(x = Time, y = weight, color = Chick)) + geom_line() + facet_wrap(~Diet, scales = 'free_y') + labs(title = "Line Plots of Weight over Time by Diet", x = "Time", y = "Weight", color = "Chick")
```
